Turbulence and Particle Acceleration in Giant Radio Halos: the Origin of Seed Electrons

### 2015-10-01 paper.tex
test

%\documentclass[prl,twocolumn,amsmath,amsfonts,amssymb,showpacs,usenatbib]{revtex4} 
%\documentclass[useAMS,usenatbib]{mn2e}
%\usepackage{epsfig,psfrag,graphicx,xspace}
%\usepackage[normalem]{ulem}
%\usepackage{txfonts}
%\usepackage{natbib}
%\usepackage{amsfonts,amssymb}
%\usepackage{amsmath,amssymb}
%\usepackage{graphicx}

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{epsfig,psfrag,txfonts,xspace}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsfonts,amsbsy,amssymb}

% --- journals --- %
\newcommand{\aap}{A{\&}A}
\newcommand{\aaps}{A{\&}AS}
\newcommand{\aapr}{AAPR}
\newcommand{\aj}{AJ}
\newcommand{\apj}{ApJ}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\apss}{Ap{\&}SS}
\newcommand{\araa}{ARA{\&}A}
\newcommand{\memsai}{{Mem.~Soc.~Astron.~Italiana}}
\newcommand{\mnras}{MNRAS}
\newcommand{\ssr}{SSRv}
\newcommand{\physrep}{Phys.~Rep.}
\newcommand{\jcomp}{J.~Comput.~Phys.}
\newcommand{\jgr}{JGR}
\newcommand{\grl}{GRL}
\newcommand{\Mstream}{{\it M-streaming}\xspace}
\newcommand{\Mflatturb}{{\it M-turbulence}\xspace}
\newcommand{\Mprimary}{{\it M-primaries}\xspace}

% --- macros --- %
\newcommand{\dd}{\mathrm{d}}
%\newcommand{\rmn}{\mathrm}
\newcommand{\Vph}{V_\mathrm{ph}}
\newcommand{\mug}{\mu G}

\def\del#1{{}}
%\def\del#1{{\bf (DELETED TEXT)}}
\def\C#1{{\bf #1}}
\def\AP#1{{\bf  AP: #1}}
\def\SPO#1{{\bf SPO: #1}}
\def\CP#1{{\bf  #1}}

@@

\author[A. Pinzke, S. Peng Oh and C. Pfrommer] 
{Anders Pinzke$^{1,2}$\thanks{apinzke@fysik.su.se (AP); peng@physics.ucsb.edu (PO); christoph.pfrommer@h-its.org (CP)}, S. Peng Oh$^{3}$ and Christoph Pfrommer$^{4}$\footnotemark[1]\\
$^{1}$The Oskar Klein Centre for Cosmoparticle Physics, Department
  of Physics, Stockholm University, AlbaNova University Center, SE - 106 91
  Stockholm, Sweden\\
$^{2}$Dark Cosmology Center, University of Copenhagen,
  Juliane Maries Vej 30, DK-2100 Copenhagen, Denmark\\
  $^{3}$University of California - Santa Barbara,
  Department of Physics, CA 93106-9530, USA\\
$^{4}$Heidelberg Institute for Theoretical Studies
  (HITS), Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany}

@@

\begin{abstract}
About $1/3$ of X-ray-luminous clusters show smooth, unpolarized radio
emission on $\sim$Mpc scales, known as giant radio halos. One
promising model for radio halos is Fermi II acceleration of seed
relativistic electrons by turbulence of the intracluster medium (ICM);
Coulomb losses prohibit acceleration from the thermal pool. However,
the origin of seed electrons has never been fully explored. Here, we
integrate the Fokker-Planck equation of the cosmic ray (CR) electron
and proton distributions in a cosmological simulations of cluster
formation. For standard assumptions, structure formation shocks lead
to a seed electron population which produces too centrally
concentrated radio emission. Instead, we present three realistic
scenarios that each can reproduce the spatially flat radio emission
observed in the Coma cluster: (1) the ratio of injected turbulent
energy density to thermal energy density increase significantly with
radius, as seen in cosmological simulations. This generates a flat
radio profile even if the seed population of CRs is steep with
radius. (2) Self-confinement of energetic CR protons can be
inefficient, and CR protons may stream at the Alfv{\'e}n speed to the
cluster outskirts when the ICM is relatively quiescent. A spatially
flat CR proton distribution develops and produces the required
population of secondary seed electrons. (3) The CR proton to electron
acceleration efficiency $K_{ep} \sim 0.1$ is assumed to be larger than
in our Galaxy ($K_{ep} \sim 10^{-2}$), due to the magnetic geometry at
the shock. The resulting primary electron population dominates. Due to
their weaker density dependence compared to secondary electrons, these
primaries can also reproduce radio observations.  These competing
non-trivial solutions provide incisive probes of non thermal processes
in the high-$\beta$ ICM.
\end{abstract}

% --- keywords --- %
\begin{keywords}
  magnetic fields, cosmic rays, radiation mechanisms: non-thermal, elementary
  particles, galaxies: cluster: general, Galaxy: fundamental parameters
\end{keywords}
%\maketitle

@@

One promising model for RHs is re-energetization of seed suprathermal
electrons by Fermi II acceleration when ICM turbulence becomes transonic during
mergers \citep{2001MNRAS.320..365B,brunetti07}. Due to the short radiative
cooling time of high-energy relativistic electrons, the cluster synchrotron
emission quickly fades away after a merger, which naturally explains the
observed bimodality of RHs \cite[see
 e.g.][]{2013MNRAS.429.3564D,2014MNRAS.443.3564D}.

However, there is a salient piece missing in the turbulent
reacceleration model. It relies heavily on the assumption of an
abundant, volume-filling population of seed suprathermal electrons;
direct Fermi II acceleration from the thermal pool is precluded by
strong Coulomb losses
\citep{2008ApJ...682..175P,2012ApJ...759..113C}. These seeds are
presumed to be either fossil CR electrons (CRes) accelerated by
diffusive shock acceleration (DSA) during structure formation
\citep{1999ApJ...520..529S}, or secondaries injected by hadronic
interaction of CR protons (CRps) with thermal protons
\citep{brunetti11}. While analytic estimates have been made, there has
been no ab initio demonstration that structure formation can lead to
the required abundance of seed electrons with the correct spatial and
spectral characteristics. This is a non-trivial requirement: Coulomb
cooling in dense cluster cores is severe, and DSA fossil electrons may
not survive. On the other hand, for secondaries to constitute the seed
population, the CRp population required in the best-studied case of
the Coma cluster must have a very broad and flat (or even slightly
inverted) spatial profile \citep{brunetti12}, in contrast with the
thermal plasma whose energy density declines steeply with radius. In
Fig.~\ref{fig:Edens} we show that such a distribution is not predicted by
cosmological simulations \cite[see
  also][]{pinzke10,2014MNRAS.439.2662V}.
\begin{figure}
  \begin{minipage}{1.0\columnwidth}
    \includegraphics[width=1.0\columnwidth]{figures/fCR.radius.coma.g72a.Rad14.2400p.z0.NL.xKR.eb23.eI066.DII.140.v5.halo.test5.eps}
    \caption{Spatial distribution of cosmic ray proton (CRp) energy
      density in Coma cluster. The red dash-dotted line shows the
      required distribution of seed CRps that generate secondary
      electrons via p-p collisions required to reproduce Coma radio
      brightness observations after Fermi-II reacceleration
      \citet{brunetti12}. The blue solid line shows the distribution of
      fossil CRps found in cosmological simulations, which disagrees
      with the required profile. We normalize the required distribution
      of CRps by fixing the total CRp energy $E_p$ to 0.3 percent of the
      total thermal energy, consistent with observations
      \citet{2014ApJ...787...18A,2012ApJ...757..123A}.}
    \label{fig:Edens}
  \end{minipage}
\end{figure}
Indeed, arriving at a seed population with the required
characteristics is highly constraining, and has the potential to teach
us much about the origin of CRps/CRes in clusters. \AP{In this paper
  we consider 3 new scenarios that individually or combined can
  reproduce the observed radio profiles and spectrum in the Coma
  cluster without violating gamma-ray constraints:} (i) Our model
\Mflatturb: a significantly spatially flatter turbulent profile than
what was adopted in \citep{brunetti12}, which allows seed CRps to
follow the steep profile that is suggested by structure formation
simulations. (ii) Our model \Mstream: streaming CRps that produce flat
distributions of CRs in the ICM
\citep{ensslin11,wiener13,2014MNRAS.438..124Z}, which also flattens
the secondary electron distribution. (iii) Our model \Mprimary: if the
acceleration efficiency of CRps is below about $0.1$~{\%} in weak
(perpendicular) shocks and the ratio of injected electrons-to-protons
$K_{\rmn{ep}} \sim 0.1$, this yields a dominant primary population
with a flat spatial distribution, since primaries have a weaker
density dependence than secondaries.

\AP{To pursue these three possibilities further, we employ cosmological
  simulations of CRs in clusters. In tandem with new insights from our
  recent work on DSA generated fossil electrons \citep{pinzke13}, we
  generate the first quantitative calculation of primary and secondary
  seed electrons and compare to observations.}

\sout{In this work we pursue these three possibilities further. We
  employ cosmological simulations of CRs in clusters, in tandem with
  new insights from our recent work on DSA generated fossil electrons
  %\citep{pinzke13}, to generate the first quantitative calculation of
  , to generate the first quantitative calculation of
  primary and secondary seed electrons.}

% --- section: Method --- %
{\bf Method.}  The transport of relativistic electrons and protons in
the ICM is a complex process that depends both on the details of the
thermal component (gas density, temperature, and pressure) as well as
non-thermal component (turbulence, magnetic fields, fossil CRs). We
use high resolution galaxy cluster simulations to derive the thermal
and fossil CR properties (shock accelerated primary CRes and CRps, as
well as secondary CRes produced in p-p collisions)
\citep{2007MNRAS.378..385P,pfrommer08,pinzke10,pinzke13}. In this
paper we focus on our simulated cluster, g72a, which is a massive
$1.6\times10^{15}\,M_\odot$ cluster that experienced a merger about
1-2 Gyrs ago \citep{2009MNRAS.399..497D}. Since the cluster mass,
density and temperature profiles are all similar to the well studied
Coma cluster \citep{2007MNRAS.378..385P,pinzke10}, we will compare our
calculations to radio and gamma-ray observations of Coma.

In our Galaxy, the CRe-to-CRp ratio at a few GeV is $K_{\rmn{ep}} \sim
10^{-2}$. Hence, we adopt this as a fiducial value for the CRe-to-CRp
acceleration efficiency (see \citep{pinzke13} for more
discussion). However, as recent PIC simulations have shown, this is
likely very different at weak shocks, with electrons efficiently
accelerated at perpendicular shocks
\citep{2014ApJ...794..153G,2014ApJ...797...47G}, and ions efficiently
accelerated at parallel shocks \citep{2014ApJ...783...91C}. Thus,
depending on magnetic geometry, $K_{\rmn{ep}}$ could be either larger
or smaller. \AP{In this work we use a simple test-particle model for
  the CRp acceleration, where we adopt a fit to Monte Carlo
  simulations of the thermal leakage process that relates the momentum
  of injected protons to the thermal energy of the shocked plasma
  \citep{kang11}. We set an upper limit on the ratio of accelerated
  proton-to-dissipated energy in the downstream of strong shocks that
  varies from 1-10\%, depending on the adopted model (for more
  details, see the Results section). In order not to violate energy
  conservation with a fixed limit for the acceleration efficiency, we
  steepen the spectral index of the accelerated CRps at the limit
  \citep{pinzke13}. For weak shocks the efficiency is a factor 10-100
  lower and well below the limit. Some observations of radio relics
  suggest high values of $K_{\rmn{ep}}$, due to the absence of
  gamma-ray emission, which probes the CRp population
  \citep{2014MNRAS.437.2291V}. This suggests primary CRes as a viable
  alternative scenario to secondary CRes as seeds for the giant
  RHs. In our {\it M-primaries} scenario, we assume that the momentum
  of injected electrons and protons are equal, and in addition we
  adopt a $K_{\rmn{ep}} =0.1$ (viable for primarily perpendicular
  shocks). While this appears to contradict the radial bias of
  magnetic fields in the bulk of the ICM as suggested by observations
  \citep{2010NatPh...6..520P} and cluster simulations
  \citep{2011ApJ...740...81R}, it does not necessarily apply to the
  accretion shock regions, which show a field geometry with a net
  perpendicular bias with respect to the shock normal---at least for
  giant radio relics \citep{2010Sci...330..347V}. We leave further
  exploration of parameter space to future work.}

As previously noted, secondaries produced by shock accelerated CRp
have the wrong spatial profile to explain RH observations; because
they arise from a two body process, they are too centrally
concentrated. They also produce $\gamma$-ray emission in excess of
Fermi-LAT upper limits
\citep{2012ApJ...757..123A,brunetti12,2014ApJ...787...18A}. However,
if CRps stream in the ICM, then their spatial profile could
potentially flatten sufficiently \citep{ensslin11,wiener13}. This
scenario is very attractive: it generates seed electrons with the
right spatial footprint, and by removing CRps from the core, obeys
gamma-ray constraints. Turbulence plays two opposing roles:
Alfv{\'e}nic turbulence damps waves generated by the CR streaming
instability \citep{yan02,farmer04}, thus reducing self-confinement;
but compressible fast modes scatter CRs directly. Turbulent damping is
still efficient for highly subsonic conditions \citep{wiener13}, while
we assume compressible fast modes only provide effective spatial
confinement during the periods of transonic, highly super-Alfv{\'e}nic
($M_{\rm A} \sim 5$) turbulence associated with mergers. Thus, CRs can
stream out when the cluster is kinematically quiescent. Furthermore,
even Alfv{\'e}nic streaming timescales are relatively short ($\sim
0.1-0.5$ Gyr; \cite{wiener13}) compared to the timescale on which the
CRp population is built up. Based on these findings, we adopt a toy
model for our \Mstream scenario in which CR streaming quickly produces
flat CRp profiles. We assume that CRs cannot
stream significantly past perpendicular $B$-fields at the accretion
shock, so that the total number of CRs is conserved. \AP{Note that the seed CRps that are
  Fermi-II reaccelerated, and produce secondary electrons that emit
  radio emission with frequencies $\gtrsim 350$~MHz, have relatively
  high energies $\gtrsim 10$~GeV and stream more efficiently than low
  energy CRs \citep{wiener13}.}

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}, leaving exploration of parameter space
to future work. We solve the isotropic, gyro-phase averaged
Fokker-Planck equation (via a Crank-Nicholson scheme) for the time
evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|{{dp}\over{dt}}\right|_{\rm C} 
+ \frac{p}{3}\left(\vec{\nabla}\cdot \vec{\upsilon}\right)\right.\right.
\nonumber\\
&+& \left.\left. \left|{{dp}\over{dt}}\right|_{\rm r}
- {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\vec{\nabla}\cdot \vec{\upsilon}\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum $p$ and time $t$ (which is
normalized such that the number density is given by
$n_{\rmn{e}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/dt=\partial/\partial
t+\vec{\upsilon}\cdot \vec{\nabla}$ is the Lagrangian derivative,
$\vec{\upsilon}$ is the gas velocity, $|dp/dt|$ represents radiative
(r) and Coulomb (C) losses, $D_{pp}$ is the momentum space diffusion
coefficient, and $Q_{\rmn{e}}$ denotes the injection rate of primary
and secondary electrons in the ICM. The $\vec{\nabla}\cdot
\vec{\upsilon}$ terms represent adiabatic gains and losses. During
post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The
simulated cluster undergoes a major merger over the last 1-2~Gyrs that
injects large turbulent eddies. After about 1~Gyr those have decayed
down to the scale needed to reaccelerate particles \AP{and the radio
  halo turns on shortly after. This naturally explains why giant radio
  halos are only seen in a fraction of all merging clusters.} In all
our calculations we assume that turbulent reacceleration is
efficiently accelerating particles for 650 Myrs and that during this
turbulent phase CR streaming and spatial diffusion can be
neglected. In our \Mstream model, CR streaming and diffusion are
incorporated separately during kinematically quiescent times that
precede the merger. As a result, flat CRp profiles are produced on
relatively short timescales ($\sim 0.1-0.5$ Gyr).

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{epsfig,psfrag,txfonts,xspace}
\usepackage[normalem]{ulem}
\usepackage{graphicx,color}
\usepackage{natbib}
\usepackage{amsfonts,amsbsy,amssymb}

% --- journals --- %
\newcommand{\aap}{A{\&}A}
\newcommand{\aaps}{A{\&}AS}
\newcommand{\aapr}{AAPR}
\newcommand{\aj}{AJ}
\newcommand{\apj}{ApJ}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\apss}{Ap{\&}SS}
\newcommand{\araa}{ARA{\&}A}
\newcommand{\memsai}{{Mem.~Soc.~Astron.~Italiana}}
\newcommand{\mnras}{MNRAS}
\newcommand{\ssr}{SSRv}
\newcommand{\physrep}{Phys.~Rep.}
\newcommand{\jcomp}{J.~Comput.~Phys.}
\newcommand{\jgr}{JGR}
\newcommand{\grl}{GRL}
\newcommand{\nat}{Nature}
\newcommand{\Mstream}{{\it M-streaming}\xspace}
\newcommand{\Mflatturb}{{\it M-turbulence}\xspace}
\newcommand{\Mprimary}{{\it M-primaries}\xspace}

% --- macros --- %
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Vph}{V_\mathrm{ph}}
\newcommand{\mug}{\mu G}

@@

\author[A. Pinzke, S. Peng Oh and C. Pfrommer] 
{Anders Pinzke$^{1,2}$\thanks{apinzke@fysik.su.se (AP); peng@physics.ucsb.edu (SPO); christoph.pfrommer@h-its.org (CP)}, S. Peng Oh$^{3}$ and Christoph Pfrommer$^{4}$\footnotemark[1]\\
$^{1}$The Oskar Klein Centre for Cosmoparticle Physics, Department
  of Physics, Stockholm University, AlbaNova University Center, SE - 106 91
  Stockholm, Sweden\\
$^{2}$Dark Cosmology Center, University of Copenhagen,
  Juliane Maries Vej 30, DK-2100 Copenhagen, Denmark\\
  $^{3}$University of California - Santa Barbara,
  Department of Physics, CA 93106-9530, USA\\
$^{4}$Heidelberg Institute for Theoretical Studies
  (HITS), Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany}

@@

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarized
  radio emission on $\sim$Mpc scales, known as giant radio halos. One
  promising model for radio halos is Fermi II acceleration of seed
  relativistic electrons by turbulence in the intracluster medium
  (ICM); Coulomb losses prohibit acceleration from the thermal
  pool. However, the origin of seed electrons has never been fully
  explored. Here, we integrate the Fokker-Planck equation of the
  cosmic ray (CR) electron and proton distributions in cosmological
  simulations of cluster formation. For standard assumptions,
  structure formation shocks lead to a seed electron population which
  produces too centrally concentrated radio emission. Instead, we
  present three plausible scenarios for the seed CRs that each
  can reproduce the spatially flat radio emission observed in the Coma
  cluster. (1) The CR proton to electron acceleration efficiency
  $K_\rmn{ep} \sim 0.1$ is assumed to be larger than in our Galaxy
  ($K_\rmn{ep} \sim 10^{-2}$), due to the magnetic geometry at the
  shock. The resulting primary electrons dominate the radio emission,
  which is more extended in comparison to radio emission from
  secondary electrons that result from hadronic CR interactions in the
  ICM. (2) CR protons may stream at the Alfv{\'e}n speed to the
  cluster outskirts when the ICM is relatively quiescent. A spatially
  flat CR proton distribution develops and produces the required
  population of secondary seed electrons. (3) The ratio of injected
  turbulent energy density to thermal energy density increase
  significantly with radius, as seen in cosmological simulations. This
  generates a flat radio profile when CRs are reaccelerated even
  if the seed population of CRs is steep with radius. These competing
  non-trivial solutions provide incisive probes of non thermal
  processes in the high-$\beta$ ICM.
\end{abstract}

%% --- keywords --- %
%\begin{keywords}
%  magnetic fields, cosmic rays, radiation mechanisms: non-thermal, elementary
%  particles, galaxies: cluster: general, Galaxy: fundamental parameters
%\end{keywords}
%%\maketitle

@@

There have been two competing models proposed to explain RHs.  The
radio emitting electrons in the ``hadronic model'' are produced in
inelastic (hadronic) CR proton interactions with protons of the
ambient thermal ICM
(\citealp{1980ApJ...239L..93D,1999APh....12..169B,2001ApJ...562..233M,
  2004A&A...413...17P,2008MNRAS.385.1211P,ensslin11}). CR protons and
heavier nuclei may have been accelerated and injected into the ICM by
structure formation shocks, active galactic nuclei and galactic
winds. However, the strong bimodality that separates X-ray luminous
clusters into radio-active and radio-quite clusters (and requires a
fast switch on/off mechanism of the RH emission) and the very extended
RH emission at low frequencies in Coma (352~MHz) represent a major
challenge to this model class \citep{brunetti12,2014MNRAS.438..124Z}.

The alternative model for RHs is re-energization of seed
suprathermal electrons by Fermi II acceleration when ICM turbulence
becomes transonic during mergers
\citep{1987A&A...182...21S,1993ApJ...406..399G,2001MNRAS.320..365B,
  2004MNRAS.350.1174B,brunetti07,brunetti11}. Due to the short
radiative cooling time of high-energy relativistic electrons, the
cluster synchrotron emission quickly fades away after a merger, which
naturally explains the observed bimodality of RHs \cite[see
  e.g.][]{2013MNRAS.429.3564D,2014MNRAS.443.3564D}.

However, there is a salient piece missing in the turbulent
reacceleration model. It relies heavily on the assumption of an
abundant, volume-filling population of seed suprathermal electrons;
direct Fermi II acceleration from the thermal pool is precluded by
strong Coulomb losses
\citep{2008ApJ...682..175P,2012ApJ...759..113C}. These seeds are
presumed to be either fossil CR electrons (CRes) accelerated by
diffusive shock acceleration (DSA) during structure formation
\citep{1999ApJ...520..529S}, or secondaries injected by hadronic
interaction of CR protons (CRps) with thermal protons
\citep{brunetti11}. While analytic estimates have been made, there has
been no ab initio demonstration that structure formation can lead to
the required abundance of seed electrons with the correct spatial and
spectral characteristics. This is a non-trivial requirement: Coulomb
cooling in dense cluster cores is severe, and DSA fossil electrons may
not survive. On the other hand, for secondaries to constitute the seed
population, the CRp population required in the best-studied case of
the Coma cluster must have a very broad and flat (or even slightly
inverted) spatial profile \citep{brunetti12}, in contrast with the
thermal plasma whose energy density declines steeply with radius. In
Fig.~\ref{fig:Edens} we show that such a distribution is not predicted by
cosmological simulations \cite[see
  also][]{pinzke10,2014MNRAS.439.2662V}.
\begin{figure*}
  \includegraphics[width=1.0\columnwidth]{figures/fCR.radius.coma.g72a.Rad14.2400p.z0.NL.xKR.eb23.eI066.DII.140.v5.halo.test5.eps}
  \includegraphics[width=1.0\columnwidth]{figures/turb_profile_ratio.eps}
  \caption{Left: spatial distribution of CRp energy density in the
    Coma cluster. The red dash-dotted line shows the required
    distribution of seed CRps that generate secondary electrons via
    p-p collisions required to reproduce Coma radio brightness
    observations after Fermi-II reacceleration \citep{brunetti12}. The
    blue solid line shows the distribution of fossil CRps found in
    cosmological simulations, which disagrees with the required
    profile. We normalize the required distribution of CRps by fixing
    the total CRp energy $E_\rmn{p}$ to 0.3 percent of the total
    thermal energy, consistent with observations
    \citep{2014ApJ...787...18A,2012ApJ...757..123A}. Right: the ratio
    of turbulent-to-thermal energy densities in our three models which
    is parametrised as $\epsilon_\rmn{turb} \propto
    \epsilon_\rmn{th}^{\alpha_{\rmn{tu}}}$. They are normalized such
      that the total turbulent energy for each scenario makes up 20\%
      of the total thermal energy inside the radio halo. Solid lines
      show the median ratio, while the dotted lines show the 68
      percentiles.}
    \label{fig:Edens}
\end{figure*}

Indeed, arriving at a seed population with the required
characteristics is highly constraining, and has the potential to teach
us much about the origin of CRps/CRes in clusters. In this paper we
consider three new scenarios that individually or combined can
reproduce the observed radio profiles and spectrum in the Coma cluster
without violating gamma-ray constraints:
\begin{enumerate}
\item {\bf Model {\em M-primaries}.} If the acceleration efficiency of CRps is below
  about $0.1$~{\%} in weak (perpendicular) shocks and the ratio of injected
  electrons-to-protons $K_{\rmn{ep}} \sim 0.1$, this yields a dominant primary
  population with a flat spatial distribution, since primaries have a weaker
  density dependence than secondaries.
\item {\bf Model {\em M-streaming}.} Here, we account for streaming CRps that
  produce flat distributions of CRps in the ICM
  \citep{ensslin11,wiener13,2014MNRAS.438..124Z}, which also flattens the
  secondary electron distribution.
\item {\bf Model {\em M-turbulence}.} Here, we adopt a significantly spatially
  flatter turbulent profile than what was adopted in \citet{brunetti12},
  but where seed CRps follow the steep profile that is suggested by
  structure formation simulations.
\end{enumerate}

To pursue these three possibilities further, we employ cosmological
simulations of CRs in clusters. In tandem with new insights from our
recent work on DSA generated fossil electrons \citep{pinzke13}, we
generate the first quantitative calculation of primary and secondary
seed electrons and compare to observations in the Coma cluster. For
the most part, we adopt parameters that are used in previous work: for
turbulence, see \citet[][ including turbulent spectrum, injection
  scale, energy density of turbulence]{brunetti07,brunetti11},
for CR acceleration in shocks, see \citet[][ in particular
  $K_{\rmn{ep}}$, acceleration efficiency]{pinzke13}, and for
cosmological simulations, see \citet[][ i.e., shock history,
  temperature profile, density profile, magnetic field
  profile]{pinzke10}. All these parameters are uncertain, and could
potentially change our predictions. We leave exploration of parameter
space and the interaction between free parameters to future work.

% --- section: Method --- %
\section{Method}

The transport of relativistic electrons and protons in
the ICM is a complex process that depends both on the details of the
thermal component (gas density, temperature, and pressure) as well as
non-thermal component (turbulence, magnetic fields, fossil CRs). We
use high resolution galaxy cluster simulations to derive the thermal
and fossil CR properties (shock accelerated primary CRes and CRps, as
well as secondary CRes produced in p-p collisions)
\citep{2007MNRAS.378..385P,pfrommer08,pinzke10,pinzke13}. In this
paper we focus on our simulated cluster, g72a, which is a massive
$1.6\times10^{15}\,M_\odot$ cluster that experienced a merger about
1-2 Gyrs ago \citep{2009MNRAS.399..497D}. Since the cluster mass,
density and temperature profiles are all similar to the well studied
Coma cluster \citep{2007MNRAS.378..385P,pinzke10}, we will compare our
calculations to radio and gamma-ray observations of Coma.

\subsection{Acceleration}

In our Galaxy, the CRe-to-CRp ratio at a few GeV is $K_{\rmn{ep}} \sim
10^{-2}$. Hence, we adopt this as a fiducial value for the CRe-to-CRp
acceleration efficiency (see \citet{pinzke13} for more
discussion). However, as recent PIC simulations have shown, this is
likely very different at weak shocks, with electrons efficiently
accelerated at perpendicular shocks
\citep{2014ApJ...794..153G,2014ApJ...797...47G}, and ions efficiently
accelerated at parallel shocks \citep{2014ApJ...783...91C}. Thus,
depending on magnetic geometry, $K_{\rmn{ep}}$ could be either larger
or smaller. In this work we use a simple test-particle model for the
CRp acceleration, where we adopt a fit to Monte Carlo simulations of
the thermal leakage process that relates the momentum of injected
protons to the thermal energy of the shocked plasma \citep{kang11}. We
set an upper limit on the ratio of accelerated proton-to-dissipated
energy in the downstream of strong shocks that varies from 1-10\%,
depending on the adopted model (for more details, see Section
\ref{sec:results}). In order not to violate energy conservation with a
fixed limit for the acceleration efficiency, we steepen the spectral
index of the accelerated CRps at the limit \citep{pinzke13}, which
mimics the non-linear CR modification of the shock
\citep{ellison00}. For weak shocks the efficiency is a factor 10-100
lower and well below the limit. Some observations of radio relics
suggest high values of $K_{\rmn{ep}}$, due to the absence of gamma-ray
emission, which probes the CRp population
\citep{2014MNRAS.437.2291V}. This suggests primary CRes as a viable
alternative scenario to secondary CRes as seeds for the giant RHs. In
our {\em M-primaries} scenario, the injected distribution of CRes is
derived in the same way as for the CRps. Once they have been
accelerated to relativistic energies, injected electrons and protons
are indistinguishable. We therefore assume that CRp and CRe have the
same distribution function $f_\rmn{CRe}(p) = K_\rmn{ep}
f_\rmn{CRp}(p)$, with a different normalization (due to differing
acceleration efficiencies) $K_\rmn{ep}=0.1$ (viable for primarily
perpendicular shocks; \citet{2014ApJ...794..153G}). While this
appears to contradict the radial bias of magnetic fields in the bulk
of the ICM as suggested by observations \citep{2010NatPh...6..520P}
and cluster simulations \citep{2011ApJ...740...81R}, it does not
necessarily apply to the accretion shock regions, which show a field
geometry with a net perpendicular bias with respect to the shock
normal---at least for giant radio relics \citep{2010Sci...330..347V}.

\subsection{Cosmic ray transport}
As previously noted, secondaries produced by shock accelerated CRp
have the wrong spatial profile to explain RH observations; because
they arise from a two body process, they are too centrally
concentrated. They also produce $\gamma$-ray emission in excess of
Fermi-LAT upper limits
\citep{2012ApJ...757..123A,brunetti12,2014ApJ...787...18A}. However,
if CRps stream in the ICM, then their spatial profile could
potentially flatten sufficiently \citep{ensslin11,wiener13}. This
scenario is very attractive: it generates seed electrons with the
right spatial footprint, and by removing CRps from the core, obeys
gamma-ray constraints. Turbulence plays two opposing roles:
Alfv{\'e}nic turbulence damps waves generated by the CR streaming
instability \citep{yan02,farmer04}, thus reducing self-confinement;
but compressible fast modes scatter CRs directly. Turbulent damping is
still efficient for highly subsonic conditions \citep{wiener13}, while
we assume compressible fast modes only provide effective spatial
confinement during the periods of transonic, highly super-Alfv{\'e}nic
($M_{\rm A} \sim 5$) turbulence associated with mergers. Thus, CRs can
stream out when the cluster is kinematically quiescent. Furthermore,
even Alfv{\'e}nic streaming timescales are relatively short
\cite[$\sim 0.1-0.5$ Gyr;][]{wiener13} compared to the timescale on
which the CRp population is built up. Based on these findings, we
adopt a toy model for our \Mstream scenario in which CR streaming
quickly produces flat CRp profiles. We assume that CRs cannot stream
significantly past perpendicular $B$-fields at the accretion shock, so
that the total number of CRs is conserved within the galaxy cluster
during the streaming process.

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|{{dp}\over{dt}}\right|_{\rm C} 
+ \frac{p}{3}\left(\vec{\nabla}\cdot \vec{\upsilon}\right)\right.\right.
\nonumber\\
&+& \left.\left. \left|{{dp}\over{dt}}\right|_{\rm r}
- {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\vec{\nabla}\cdot \vec{\upsilon}\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum $p$ and time $t$ (which is
normalized such that the number density is given by
$n_{\rmn{e}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/dt=\partial/\partial
t+\vec{\upsilon}\cdot \vec{\nabla}$ is the Lagrangian derivative,
$\vec{\upsilon}$ is the gas velocity, $|dp/dt|$ represents radiative
(r) and Coulomb (C) losses, $D_{pp}$ is the momentum space diffusion
coefficient, and $Q_{\rmn{e}}$ denotes the injection rate of primary
and secondary electrons in the ICM. The $\vec{\nabla}\cdot
\vec{\upsilon}$ terms represent adiabatic gains and losses. During
post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The
simulated cluster undergoes a major merger over the last 1-2~Gyrs that
injects large turbulent eddies. After about 1~Gyr those have decayed
down to the scale needed to reaccelerate particles and the radio halo
turns on shortly after. This delay naturally explains why giant radio
halos are only seen in a fraction of all merging clusters. In all our
calculations we assume that turbulent reacceleration efficiently
accelerates particles for 650 Myrs (which is roughly the cascade time
on which turbulence is damped) and that during this turbulent phase CR
streaming and spatial diffusion can be neglected. In our \Mstream
model, CR streaming and diffusion are incorporated separately during
kinematically quiescent times that precede the merger. As a result,
flat CRp profiles are produced on relatively short timescales ($\sim
0.1-0.5$ Gyr).
